---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  error = TRUE
)
```

# aPPR

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/aPPR)](https://CRAN.R-project.org/package=aPPR)
<!-- badges: end -->

`aPPR` helps you calculate approximate personalized pageranks from large graphs, including those that can only be queried via an API. `aPPR` additionally performs degree correction and regularization, allowing users to recover blocks from stochastic blockmodels.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RoheLab/aPPR")
```

## Find the personalized pagerank of a node in an `igraph` graph

This is a basic example which shows you how to solve a common problem:

```{r igraph-example}
library(aPPR)
library(igraph)

set.seed(27)

graph2 <- sample_pa(100)

appr(graph2, seeds = "5", verbose = TRUE)
```

## Find the personalized pagerank of a Twitter user using `rtweet`

```{r rtweet-example, eval = FALSE}
graph <- rtweet_graph()

batch_appr(graph, "fchen365", epsilon = 1e-4, verbose = TRUE)
```

## Limitations

- Connected graph assumption, what results look like when we violate this assumption
- Sampling is one node at a time

## Speed ideas

compute is not an issue relative to actually getting data

Compute time ~ access from Ram time << access from disk time << access from network time.

Make requests to API in bulk, memoize everything, cache / write to disk in a separate process?

General pattern: cache on disk, and also in RAM

# References

1. [Targeted sampling from massive Blockmodel graphs with personalized PageRank][chen]

2. [Local Graph Partitioning using PageRank Vectors][andersen]

[chen]: https://arxiv.org/abs/1910.12937
[andersen]: http://www.leonidzhukov.net/hse/2015/networks/papers/andersen06localgraph.pdf
