---
title: "Extending aPPR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending aPPR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Suppose you want to calculate personalized PageRanks for some graph that is not supported by `aPPR`. You can extend `aPPR` to arbitrary graph objects, so long the graph object is an S3 object with methods:

- `check()`
- `node_degrees()`
- `neighborhood()`
- `appr()` (optional)

See the documentation for those S3 generics to understand the generic specification fully!

We demonstrate how to implement these methods below for a new, custom graph object. In this case, we will consider the PubMed citation network, which we will interact via the PubMed API, using the `rentrez` package. First, we define a constructor function that returns a custom S3 graph object that subclasses `abstract_graph`. You can install rentrez with `pak::pak("ropensci/rentrez")`, and do not need to set up any authentication to begin using the API.

```{r}
library(aPPR)
library(rentrez)

# Constructor for PubMed graph object (defined over API)
pubmed_graph <- function(max_attempts = 3) {
  
  if (!requireNamespace("rentrez", quietly = TRUE)) {
    stop(
      "`rentrez` package must be installed to use `pubmed_graph()`",
      call. = FALSE
    )
  }
  
  # could make this into an argument if you wanted to let the user have
  # control over this
  citation_types_of_interest <- c(
    "pubmed_pubmed",
    "pubmed_pubmed_alsoviewed",
    "pubmed_pubmed_citedin",
    "pubmed_pubmed_refs",
    "pubmed_pubmed_reviews"
  )

  agraph <- abstract_graph(
    subclass = "pubmed",
    citation_types_of_interest = citation_types_of_interest,
    max_attempts = max_attempts
  )
  
  agraph
}

graph <- pubmed_graph()
graph
```

Now we want to implement S3 methods for the `pubmed` object. We'll want to test these, which we can do using a seed node, in this case PubMed paper 30345262.

```{r}
#' Check method for `pubmed` graph objects
#'
#' @param graph A `pubmed` graph object
#' @param nodes A **character** vector of node ids. **Can be empty!**
#'
#' @return A **character** vector of node ids that we can reach in the graph.
#'   For example, some nodes ids may not be reachable due to API failures,
#'   or, more generally, permissions failures.
#'
#'   If `nodes` is the empty vector, returns the empty vector. Be sure to
#'   handle this edge case.
check.pubmed <- function(graph, nodes) {

  logger::log_debug(glue::glue("Checking nodes"))
  logger::log_trace(glue::glue("Checking nodes: {nodes}"))

  # handle the edge case where no nodes are passed
  if (length(nodes) < 1)
    return(character(0))
  
  # TODO: check if we can get node degrees for each node in `nodes`
  # TODO: check if we can get the neighbor for each node in `nodes`
  # ideally we check this in some simpler faster way than 
  # pulling all of that information, normally we just want to check if an
  # item is in the database
  
  # if you're API is consistently reliable and there are no permission issues
  # you could conceivably not do any actual checking here to save on API
  # calls
  
  logger::log_debug(glue::glue("Done checking nodes"))

  # TODO
  nodes
}
```

Now we test our implementation. To do this, we should give at least one good node id, and at least one bad node id. Only the good node id should be returned.

```{r}
good_node_ids <- c("30345262", "29624432", "29867837")
bad_node_id <- "I am a pumpkin"

check(graph, c(good_node_ids, bad_node_id))
```

Now we move on to the node degree method.

```{r}
#' Degree method for `pubmed` graph objects
#'
#' @param graph A `pubmed` graph object
#' @param nodes A **character** vector of node ids. **Cannot be empty.** Should
#'   not contain duplicates if `check()` is properly implemented and does
#'   not output duplicates.
#'
#' @return TODO
node_degrees.pubmed <- function(graph, nodes) {

  logger::log_debug(glue::glue("Getting node degrees"))
  logger::log_trace(glue::glue("Getting node degrees for node: {nodes}"))

  # assumes that you want any errors / empty rows when accessing this
  # data, i.e. that the nodes have already been checked

  # at this point, all the nodes in `nodes` have been checked, so we should
  # in theory be able to get node degrees for them
  
  # TODO

  logger::log_debug(glue::glue("Done getting node degrees"))

  # if your graph is symmetric, set both in_degree and out_degree to the same
  # thing
  
  list(
    in_degree = NULL,
    out_degree = NULL
  )
}
```

To test this method, we should pass a character vector of several good node ids.

```{r}
node_degrees(graph, good_node_ids)
```

```{r}
#' Neighborhood method for `pubmed` graph objects
#'
#' @param graph A `pubmed` graph object
#' @param nodes A length one character vector, for a node in the graph with
#'   at least one outgoing edge.
#'
#' @return A **character** vector of node ids for the graph neighborhood.
#'   Should be a vector of length at least one (if the `check()` method was
#'   implemented correctly), and should not contain duplicates.
neighborhood.pubmed <- function(graph, node) {

  logger::log_debug(glue::glue("Getting neighborhood: {node}"))
  
  for (i in 1:graph$max_attempts) {
    logger::log_trace(glue::glue("Attempt {i}/{graph$max_attempts} to get neighborhood of: {node}"))
    tryCatch({
      cites <- entrez_link(dbfrom = "pubmed", db = "all", id = node)
      break
    }, error = function(e){
      if (i == graph$max_attempts){
        logger::log_debug(glue::glue("Maximum attempts to find neighborhood met, could not find: {node}"))
      }
    })
  }
  
  # i *think* this function can also return an empty character vector
  # if strictly necessary, in case the API request to get the neighborhood
  # fails, but i am not sure
  
  # TODO: i don't know what happens if loops are allowed? like, don't self cite
  
  unique(unlist(cites$link[graph$citation_types_of_interest]))
}
```

```{r}
neighborhood(graph, good_node_ids[1])
```

Lastly, you can optionally implement an `appr` method for your abstract graph subclass. In the `appr` method for the subclass, you can do things like:

- Add functionality to convert a convenient seed node name into the internal node name representation (see `appr.rtweet_graph()` for an example of this)
- Checks that you have appropriate authorization to pull information about the seed nodes
- Etc, etc

This custom subclass method will run before the general `appr.abstract_graph()`. Be sure to end the method with a call to `NextMethod()`.

```{r}
appr.pubmed <- function(graph, seeds, ...) {

  # TODO: do we want to do anything fancy here?

  NextMethod()
}
```


### Debugging

If you are accessing a graph over an API, it's likely that you will encounter edge cases where the API returns no data, or data in a format that you did not expect. We highly recommend using logging to debug your implementation when this happens, using the `logger` library. See that `logger` documentation for details. 

Find any errors, fix, and rinse and repeat until you've completed the likely unpleasant task of tracking down all the edges cases in the API.

```{r}
library(logger)

# hide basically all messages (not recommended)

# set logging threshold for code you just wrote
log_threshold(DEBUG)

# set logging threshold for aPPR package functions
log_threshold(DEBUG, namespace = "aPPR") 

appr(
  graph,                    # the graph to work with
  seeds = good_node_ids[1], # name of seed node (character)
  epsilon = 0.0005          # desired approximation quality (see ?appr)
)
```


